MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-07-11 latex/paper/sections/10-placement.tex
snapshot discussion

In force directed placers, there is a concept of global placement followed by detailed placement.
Global placement places objects freely on an analog plane and is driven purely by directed movement and analytical solution.
In global placement stage, only non-hardware related objectives are considered. 
Objects do not have to adhere to any hardware constraints and can even overlap with one another.
Afterwards, detailed placement "legalizes" the global placement by discretizing the analog positions of the objects onto the discrete grid of resources of the \texttt{device}.
Overlapping objects are iteratively spread out until every \texttt{SiteInst}-\texttt{Cell} is assigned its own unique \texttt{Site}-\texttt{BEL}.
The spreading process somewhat resembles primitive fluid dynamics simulators or thermal diffusion simulators, where packets of energy are spread through a medium until each voxel or pixel of the medium falls below a maximum threshold of energy. 
There also exists other strategies to mix packing, placement, and legalization in different orders. 
In this context, our placer follows a very straightforward pack-legalize-place strategy with SA.

>>>>

In force directed placers, there is a concept of global placement followed by detailed placement.
Global placement places object instances freely on an analog plane and is driven purely by directed movement and analytical solution.
In global placement stage, only non-hardware related objectives are considered. 
Object instance placements do not have to adhere to any hardware constraints and can even overlap with one another.
Afterwards, detailed placement "legalizes" the global placement by discretizing the analog positions of the instances onto the discrete grid of resources of the \texttt{device}.
Overlapping instances are iteratively spread out until every \texttt{SiteInst}-\texttt{Cell} is assigned its own unique \texttt{Site}-\texttt{BEL}.
The spreading process somewhat resembles primitive fluid dynamics simulators or thermal diffusion simulators, where packets of energy are spread through a medium until each voxel or pixel of the medium falls below a maximum threshold of energy. 
There also exists other strategies to mix packing, placement, and legalization in different orders. 
In this context, our placer follows a very straightforward pack-legalize-place strategy with SA.

### 2025-07-11 latex/paper/sections/11-placement_results.tex
snapshot discussion

Next is \texttt{PlacerGreedyRandom} which is represented by the purple curve.
We can observe a much steeper initial drop in system cost than the vanilla SA placer due to greedy movement acceptance.
Interestingly, this placer also settled at a lower final system cost of 333K, which is a slight 4\% improvement over vanilla SA.
We originally expected the greediness to be a detriment to finding the global minimum, but due to the inherent randomness of the algorithm, these expectations are never certain.
Broader trends will require many more test trials to observe reliably.

@@

>>>>

Next is \texttt{PlacerGreedyRandom} which is represented by the purple curve.
We can observe a much steeper initial drop in system cost than the vanilla SA placer due to greedy movement acceptance.
Interestingly, this placer also settled at a lower final system cost of 333K, which is a slight 4\% improvement over vanilla SA.
We originally expected the greediness to be a detriment to finding the global minimum, but due to the inherent randomness of the algorithm, these broader trends will require many more test trials to reliably observe.

@@

    Figures \ref{fig:PGRSnapshots}, \ref{fig:PGMSnapshots}, \ref{fig:PARSnapshots}, \ref{fig:PAMSnapshots}, and \ref{fig:PAHSnapshots} show snapshots of the placement progress on the physical \texttt{device} after 0, 10, 100, and 300 passes for each placer variant.

    Please zoom into the document to render the figures in greater detail.
    All black areas represent purely routing \texttt{Site}s that contain no logic resources. 
    All gray or dull-colored boxes represent \texttt{Site}s that are unoccupied by \texttt{SiteInst}s, while brightly filled boxes represent \texttt{Site}s that are occupied by the supported library of \texttt{SiteInst}s.
    Cyan boxes represent occupied \texttt{SLICEL}s, magenta boxes represent occupied \texttt{RAMB18E1}s, and yellow boxes represent occupied \texttt{DSP48E1}s.

    The tangle of red lines represent the nets between the \texttt{Site}s, which are yet to be routed at this stage. 
    Brighter red lines represent nets of higher total HPWL, while duller red lines represent nets of lower total HPWL.

    Notice there are two brightly red nets in every figure that have a particularly high fan-out:
    one sourced from the near-center, and another sourced from the top right of the device.

    The singular white box in the near-center of all of the figures represents an occupied \texttt{BUFGCTRL} \texttt{Site}, which is a clock buffer used to route clock trees. 
    Since clock nets tend to have very high fanout, they can use up a significant portion of the routing resource if they are routed through general routing.
    Modern FPGAs like the 7-Series have dedicated clock routes which can efficiently distribute clock signals to all logical \texttt{Site}s without burdening the general routing resources.
    These special clock routes are accessed by routing the off-chip crystal oscillator clock to one of these clock buffers which connects to the clock tree network.
    It is always best practice to use them to improve the routability of the design, and simultaneously, to reduce clock skew and improve timing. 
    Our simple FIR Filter design only operates on a single clock domain, so only one clock buffer \texttt{Site} is used. 
    An HDL design operating on many clock domains will use more clock buffers \texttt{Site}s or a dedicated clock generator such as a phase locked loop (PLL) \texttt{Site} or mixed-mode clock manager (MCMM) \texttt{Site}.

    The right and left edges of the FPGA device is where most of the general purpose IO (GPIO) is located. 
    In the top right of the device is where we source our reset signal. 
    Like the clock net, the reset net also has high fanout and is supplied to all synchronous resources.
    Unlike clock signals, reset signals are not subject to sensitive timing constraints as they are not active on most clock cycles and can tolerate higher skew.
    Thus, they are simply routed via general routing.

    Observe how \texttt{PlacerGreedyMidpoint} \ref{fig:PGMSnapshots} and \texttt{PlacerAnnealMidpoint} \ref{fig:PAMSnapshots} both show almost not shrinkage in profile after the first 10 iterations, likely owing to the inability to hill-climb.
